### 2023-01-15 tensor_clustering/ieee_revision/response_s2.tex

   Hence, we have 
   \begin{align}
       & \inf_{\hat z_1(1)} \{ \bbP[\hat z_1(1) = 1| z_1^*(1) = 2, z^*_k,  \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(1) = 2| z_1^*(1) = 1, z^*_k,  \mtheta_k^*, \tS^*] \} \\
       & \quad = 2 \bbP[ \hat z_{LS} = 1 |  z_1^*(1) = 2, z^*_k,  \mtheta_k^*, \tS^* ] \\
      & \quad  = 2 \bbP[ \onormSize{}{\my - \theta^* \mx_1^*}_F^2 \leq \onormSize{}{\my - \theta^* \mx_2^*}_F^2 | z_1^*(1) = 2, z^*_k , \mtheta_k^*, \tS^*   ] \\
      & \quad = 2 \bbP \off{ 2 \ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \geq  \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F^2 }\\
       & \quad =  2 \bbP[ N(0,1) \geq \theta^* \onormSize{}{\mx_1^* - \mx_2^*}_F /(2\sigma) ] \\
       & \quad \geq 2 \bbP[ N(0,1) \geq c_4 ]   \geq c,
   \end{align}

where the first equation holds by symmetry, the third equation holds by rearrangement, the fourth equation holds by the fact that $\ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \sim N(0, \sigma \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F)$, and $c$ is some positive constant in the last inequality.
   %for some positive constant $c$, where the calculation for the last equation follows from the calculation in equation (102) in \cite{han2020exact}.

   Plugging the inequality~\eqref{eq:inf} into the inequality~\eqref{eq:inf_lower} for all $i \in T_1^c$, then, we have 
   \begin{equation}
       \liminf_{p \rightarrow \infty}  \inf_{\hat z_1 } \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right]  \geq \liminf_{p \rightarrow \infty} \frac{ C cp }{r^3} \geq C c,
   \end{equation}
   where the last inequality follows by the condition $r = o(p^{1/3})$. By the discrete nature of the misclassification error, we obtain our conclusion
   \begin{equation}
       \liminf_{p \rightarrow \infty} \inf_{\tS^* \in  \tP_{\tS}(\gamma)}  \inf_{\hat z_{\text{stat}} }\sup_{ (z^*, \mtheta^*) \in \tP_{z, \mtheta}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]  \geq 1. 
   \end{equation}
"
   \end{proof}
    \end{quote}

    % In fact, our proof idea of minimax lower bound statements (impossibility in Theorems~\ref{thm:stats} and \ref{thm:comp}) shares the same spirit with reviewer's comment. We briefly describe the proof idea of impossibility below:

    % Let $\tP_z$ denote the space for all possible (polynomial-time) community assignment estimation $\hat z$ and $\tP(\gamma) = \tP \cap \{ \Delta_{\rm ang}^2/\sigma^2 = p^{\gamma}\}$ denote the dTBM parameter space with SNR condition. We prove the minimax lower bound by constructing fixed true parameters $(z^*, \mtheta^*, \tS^*) \in \tP(\gamma)$ such that 
    % \begin{equation}\label{eq:impossible_prove}
    %     \inf_{\hat z \in \tP_z} \sup_{(z, \mtheta, \tS) \in \tP(\gamma)} \bbE[p \ell(\hat z, z)] \geq  \inf_{\hat z \in \tP_z} \bbE[p \ell(\hat z, z^*)| (z^*, \mtheta^*, \tS^*)] \geq 1.
    % \end{equation}
    % Here are two details differ with reviewer's comment. First, the SNR condition in $\tP(\gamma)$ is $\{ \Delta_{\rm ang}^2/\sigma^2 = p^{\gamma}\}$ rather than $\{ \Delta_{\rm ang}^2/\sigma^2 \leq p^{\gamma}\}$; otherwise, for any $\gamma \leq 0$, we can always take $\tS^* \in \{\Delta_{\rm ang}^2 = 0\}$ to make dTBM non-identifiable and thereof theoretically impossible. Second, we consider one particular core tensor $\tS^*$ rather than the random tensor with entries $\tS_{ijk} = 1 + \epsilon_{ijk}, \epsilon_{ijk} \sim N(0, \delta^2)$ for some $\delta \rightarrow 0$ as $p \rightarrow \infty$. The expectation in minimax error~\eqref{eq:impossible_prove} is taken with respect to the noise in observation, rather than the randomness in the true parameters.

    % The construction of $(z^*, \mtheta^*, \tS^*) \in \tP(\gamma)$ is the key step in the proof. With a good construction, the second inequality in~\eqref{eq:impossible_prove} can be easily verified with prior works. For statistical lower bound, we take one $(z^*, \mtheta^*, \tS^*)$ with $\mtheta^* = \mathbf{1}$ and $\Delta_{\rm Euc}^2(\tS^*) = p^{\gamma}$; for computational lower bound, we construct a rank-2 $\tS^*$ to relate the HPC conjecture. The inequality~\eqref{eq:impossible_prove} then holds with \citet[Proof of Theorems 6 and 7]{han2020exact}.

    % From the perspective of proof, we agree that the minimax lower bound statements can be obtained by exploiting prior results via proof idea~\eqref{eq:impossible_prove}. However, we argue that \textbf{none of our lower bound results are corollaries of \cite{han2020exact} and dTBM-specific techniques are required.} More importantly, we want to emphasize that the proof of dTBM achievability is more difficult than that for TBM due to the extra heterogeneity.

    % In the revision, we have proved a stronger lower bound statement for statistical impossibility as following comment suggested and added more discussion in Section~\ref{sec:tbm}.

    % Define $\tP_z \coloneqq \offf{ z:  c_1 p/r \leq |z^{-1}(a)| \leq c_2 p/r, a \in [r]}$, where $c_1, c_2$ are the positive constants in dTBM parameter space~\eqref{eq:family}. The space $\tP_z$ collects all possible community assignment $z$ under the dTBM or TBM. Let $\tP_{\rm dTBM}(\gamma) \coloneqq \tP_{\rm dTBM} \cap \{\Delta_{\rm ang}^2 = p^{\gamma}\}$ refer to the space $\tP(\gamma)$ in our main text. We prove the minimax lower bound by constructing fixed true parameters $(z^*, \mtheta^*, \tS^*) \in \tP_{\rm dTBM}(\gamma)$ such that 
    % \begin{equation}
    %     \inf_{\hat z \in \tP_z} \sup_{(z, \mtheta, \tS) \in \tP_{\rm dTBM}(\gamma)} \bbE[p \ell(\hat z, z)] \geq  \inf_{\hat z \in \tP_z} \bbE[p \ell(\hat z, z^*)| (z^*, \mtheta^*, \tS^*)] \geq 1.
    % \end{equation}

    % Let $\tP_{\rm TBM}(\gamma) \coloneqq \tP_{\rm TBM} \cap \{\Delta_{\rm Euc}^2 = p^{\gamma}\}$ and $\tP_{\rm dTBM}(\gamma) \coloneqq \tP_{\rm dTBM} \cap \{\Delta_{\rm ang}^2 = p^{\gamma}\}$ denote the parameter space for TBM and dTBM with SNR requirement, where $\tP_{\rm TBM}(\gamma)$ refers to the equation (26) in \cite{han2020exact} and $\tP_{\rm dTBM}(\gamma)$ refers to the space $\tP(\gamma)$ in our main text.

%     \textbf{Response:} We thank reviewer for the suggestions. Inspired by the reviewer, we have proved a stronger statement for statistical impossibility. The new statement suggests the dTBM statistical impossibility whenever the core tensor $\tS$ leads to a week SNR with exponent $\gamma < -(K-1)$. The proof of our new statement relies on special construction of heterogeneity $\mtheta^*$ in the worst case, and previous conclusions for TBM is non-applicable under the construction. We attach our new statement in Theorem~\ref{thm:stats}, Section~\ref{sec:limits}:

%     For self-consistency, we attach the proof of the new impossibility result in Section~\ref{sec:statprove1}:

%     \begin{quote}
%       ``
%       \begin{proof}[Proof of Theorem~\ref{thm:stats} (Impossibility)]Consider the general asymetric dTBM~\eqref{eq:general_dtbm} in the special case that $p_k = p$ and $r_k = r$ for all $ k\in [K]$. For simplicity, we show the minimax rate for the estimation on the first mode $\hat z_1$; the proof for other modes are essentially the same.

%   To prove the minimax rate~\eqref{eq:minminmax}, it suffices to take an arbitrary $\tS^* \in  \tP_{\tS}(\gamma)$ wih $\gamma < -(K-1)$ and construct $(z^*_k, \mtheta^*_k)$ such that 
%   \begin{equation}
%       \inf_{\hat z_1} \bbE \left[ p\ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right]\geq 1.
%   \end{equation}

%   We first define a subset of indices $T_k \subset [p_k], k \in [K]$ in order to avoid the complication of label permutation. Based on \citet[Proof of Theorem 6]{han2020exact}, we consider the restricted family of $\hat z_k$'s for which the following three conditions are satisfied:
%   \begin{align}
%         &\text{(a)}\ \hat z_k(i)=z_k(i) \text{ for all }i\in T_k; \quad \text{(b)} \ |T^c_k|\asymp {p\over r}; \\
%         &\text{(c)}\ \min_{\pi\in \Pi}\sum_{i\in[p]}\ind\{\hat z_k(i) \neq \pi\circ z_k (i)\} = \sum_{i\in[p]}\ind\{\hat z_k(i) \neq  z_k (i)\},
%   \end{align}
% for all $k \in [K]$.
%   Now, we consider the construction:
%   \begin{enumerate}
%       \item[(i)] $\{z_k^*\}$ satisfies properties (a)-(c) with misclassification sets $T_k^c$ for all $k \in [K]$;
%       \item [(ii)] $\{\mtheta_k^*\}$ such that $\mtheta_k^*(i) \leq \sigma r^{(K-1)/2} p^{-(K-1)/2}$ for all $i \in T_k^c, k \in [K]$ and $\max_{k \in [K], a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}^2_2 \asymp p/r$.
%   \end{enumerate}
%   % There exists a $(z^*, \mtheta^*) \in \tP_{z, \mtheta}$ satisfying the property (i) and (ii) when $r < p$; otherwise, when $r = p$, we must have $\mtheta^*(i) = 1$ due to the $\ell_1$ norm constraint and we can not satisfy the property (ii).

%   Then, following the proof of \cite{gao2018community}, we have 
%   \begin{align}
%         & \inf_{\hat z_1 } \bbE \left[ \ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right] \geq  \\
%         &\frac{C}{r^3 |T_1^c|} \sum_{i \in T_1^c} \inf_{\hat z_1  (i)} \{ \bbP[\hat z_1(i) = 1| z_1^*(i) = 2, z_k^*, \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(i) = 2| z_1^*(i) = 1, z_k^*, \mtheta_k^*, \tS^*] \},
%   \end{align}
%   where $C$ is some positive constant,  $\hat z_1$ on the left hand side denote the generic assignment functions in $\tP(\gamma)$, and the infimum on the right hand side is taken over the generic assignment function family of $\hat z_1(i)$ for all node $i \in T_1^c$. Here, the factor $r^3=r\cdot r^2$ in~\eqref{eq:inf_lower} comes from two sources: $r^2\asymp {r\choose 2}$ comes from the multiple testing burden for all pairwise comparisons among $r$ clusters; and another $r$ comes from the number of elements $|T^c_k|\asymp p/r$ to be clustered.

%   Next, we need to find the lower bound of the rightmost side in~\eqref{eq:inf_lower}. 
%   %For simplicity, we show the bound for the mode-1 case $k = 1$ only. We drop the subscripts $1$ in $z_1, T_1, \mS_1, \mtheta_1$ and omit the repeated procedures for the cases of $k = 2,\ldots, K$. 
% We consider the hypothesis test based on model~\eqref{eq:general_dtbm}. First, we reparameterize the model under the construction (i)-(ii).
% \begin{equation}
%     \mx_a^* = \off{\mat_1 \of{ \tS^*\times_2 \mTheta_2^* \mM_2^*\times_3\cdots \times_K \mTheta^* \mM_K^* }}_{a:}, \quad \text{for all} \ a \in [r],
% \end{equation}
% where $\mx_a^*$'s are centroids in $\bbR^{p^{K-1}}$. Without loss of generality, we consider the lower bound for the summand in~\eqref{eq:inf_lower} for $i=1$. The analysis for other $i\in T^c_1$ are similar. For notational simplicity, we suppress the subscript $i$ and write $\my, \theta^*, z$ in place of $\my_1, \mtheta_1^*(1)$ and $z_1(1)$, respectively. The equivalent vector problem for assessing the summand in~\eqref{eq:inf_lower} is
% \begin{equation}
% \my=\theta^* \mx_{z}^*+\me,
% \end{equation}
% where $z\in \{1,2\}$ is an unknown parameter, $\theta^* \in \bbR_+$ is the given heterogeneity degree, $\mx_1^*,\mx_2^*\in\bbR^{p^{K-1}}$ are given centroids, and $\me\in\bbR^{p^{K-1}}$ consists of i.i.d.\ $N(0,\sigma^2)$ entries.  Then, we consider the hypothesis testing under the model~\eqref{eq:z}:
% \begin{equation}
%  H_0: z = 1 \text{ and } \my = \theta^* \mx_1^* + \me \quad \leftrightarrow \quad H_1: z = 2 \text{ and } \my = \theta^* \mx_2^* + \me,
% \end{equation}

% %   For an arbitrary $i \in T_1^c$, the right hand side in \eqref{eq:inf_lower} corresponds to the following hypothesis test:
% %   \begin{equation}\label{eq:hypo}
% %       H_0: \my = \mtheta_1^*(i) \mx_1^* + \me \quad \leftrightarrow \quad H_1: \my = \mtheta_1^*(i) \mx_2^* + \me,
% %   \end{equation}
% %   where $\my = [\mat_1(\tY)]_{i:}$, $\mx_a^* = [\mat_1( \tS^* \times_2 \mTheta^* \mM_2^* \times_3 \cdots \times_K \mTheta^* \mM_K^*)]_{a:}$, and $\me$ consists of i.i.d.\ $N(0,\sigma^2)$ entries. 
%   The hypothesis test~\eqref{eq:test} is a simple versus simple testing, since the assignment $z$ is the only unknown parameter in the test. 
%   By Neyman-Pearson lemma, the likelihood ratio test is optimal with minimal Type I + II error. Under Gaussian model, the likelihood ratio test is equivalent to the least square estimator $\hat z_{LS} = \argmin_{a = \{1,2\}} \onormSize{}{\my - \theta^* \mx_a^*}_F^2$.

%   Let $\mS = \mat_1(\tS)$. Note that 
%   \begin{align}
%       \onormSize{}{ \theta^* \mx_1^*  - \theta^* \mx_2^*}_F &\leq  \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\max}(\mTheta_k^*\mM_k^*)  \\
%       & \leq \theta^* \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F   \max_{k \in [K]/\{1\}, a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}_2^{K-1} \\
%       & \leq  \sigma r^{(K-1)/2} p^{-(K-1)/2} 2 c_4 p^{(K-1)/2} r^{-(K-1)/2}\\
%       & \leq 2 c_4 \sigma, 
%   \end{align}
% %   \begin{align}
% %       \onormSize{}{ \mtheta_1^*(i) \mx_1^*  - \mtheta_1^*(i) \mx_2^*}_F &\leq  \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\max}(\mTheta_k^*\mM_k^*)  \\
% %       & \leq \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F   \max_{k \in [K]/\{1\}, a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}_2^{K-1} \\
% %       & \leq  \sigma r^{(K-1)/2} p^{-(K-1)/2} 2 c_4 p^{(K-1)/2} r^{-(K-1)/2}\\
% %       & \leq 2 c_4 \sigma, \label{eq:bound}
% %   \end{align}
%   where the second inequality follows Lemma~\ref{lem:singular_thetam} and the third inequality follows from property (ii) and the boundedness constraint in $\tP_{\tS}(\gamma)$ such that $\onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F  \leq \onormSize{}{\mS_{1:}^*}_F + \onormSize{}{\mS_{2:}^*}_F \leq 2c_4$.

### 2023-01-15 tensor_clustering/ieee_revision/response_s2.tex

%   Hence, we have 
%   \begin{align}
%       & \inf_{\hat z_1(1)} \{ \bbP[\hat z_1(1) = 1| z_1^*(1) = 2, z^*_k,  \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(1) = 2| z_1^*(1) = 1, z^*_k,  \mtheta_k^*, \tS^*] \} \\
%       & \quad = 2 \bbP[ \hat z_{LS} = 1 |  z_1^*(1) = 2, z^*_k,  \mtheta_k^*, \tS^* ] \\
%       & \quad  = 2 \bbP[ \onormSize{}{\my - \theta^* \mx_1^*}_F^2 \leq \onormSize{}{\my - \theta^* \mx_2^*}_F^2 | z_1^*(1) = 2, z^*_k , \mtheta_k^*, \tS^*   ] \\
%       & \quad = 2 \bbP \off{ 2 \ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \geq  \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F^2 }\\
%       & \quad =  2 \bbP[ N(0,1) \geq \theta^* \onormSize{}{\mx_1^* - \mx_2^*}_F /(2\sigma) ] \\
%       & \quad \geq 2 \bbP[ N(0,1) \geq c_4 ]   \geq c,
%   \end{align}
% %   \begin{align}
% %       & \inf_{\hat z_1(1)} \{ \bbP[\hat z_1(1) = 1| z_1^*(i) = 2, z^*_k,  \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(1) = 2| z_1^*(i) = 1, z^*_k,  \mtheta_k^*, \tS^*] \} \\
% %       & \quad = 2 \bbP[ \hat z_{LS}(i) = 1 |  z_1^*(i) = 2, z^*_k,  \mtheta_k^*, \tS^* ] \\
% %       & \quad  = 2 \bbP[ \onormSize{}{\my - \mtheta_1(i)^* \mx_1^*}_F^2 \leq \onormSize{}{\my - \mtheta_1(i)^* \mx_2^*}_F^2 | z_1^*(i) = 2, z^*_k , \mtheta_k^*, \tS^*   ] \\
% %       & \quad =  2 \bbP[ N(0,1) \geq \mtheta_1(i)^* \onormSize{}{\mx_1^* - \mx_2^*}_F /\sigma ] \\
% %       & \quad \geq 2 \bbP[ N(0,1) \geq 2 c_4 ]   \geq c, \label{eq:inf}
% %   \end{align}
% where the first equation holds by symmetry, the third equation holds by rearrangement, the fourth equation holds by the fact that $\ang{\me, \theta^* \mx_1^* - \theta^*\mx_2^* } \sim N(0, \sigma \onormSize{}{\theta^* \mx_1^* - \theta^*\mx_2^*}_F)$, and $c$ is some positive constant in the last inequality.
%   %for some positive constant $c$, where the calculation for the last equation follows from the calculation in equation (102) in \cite{han2020exact}.

%   Plugging the inequality~\eqref{eq:inf} to inequality~\eqref{eq:inf_lower}, then, we have 
%   \begin{equation}
%       \liminf_{p \rightarrow \infty}  \inf_{\hat z_1 } \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right]  \geq \liminf_{p \rightarrow \infty} \frac{ C cp }{r^3} \geq C c,
%   \end{equation}
%   where the last inequality follows by the condition $r = o(p^{1/3})$. By the discrete nature of the misclassification error, we have 
%   \begin{equation}
%       \liminf_{p \rightarrow \infty} \inf_{\tS^* \in  \tP_{\tS}(\gamma)}  \inf_{\hat z_{\text{stat}} }\sup_{ (z^*, \mtheta^*) \in \tP_{z, \mtheta}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]  \geq 1. 
%   \end{equation}
% %   for any $r < p$ and $p$,  we have 
% %   \begin{equation}
% %         \inf_{\hat z_1 \in \tP_z} \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right] \geq \frac{ cp }{r^3} > 0,
% %   \end{equation}
% %   which indicates 
% %   \begin{equation}
% %         \inf_{\hat z_1 \in \tP_z} \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right] \geq 1, \quad \text{for every } p
% %   \end{equation}
% %   by the discrete nature of $p\ell(\hat z, z^*)$.
%   \end{proof} "
%     \end{quote}

    \item \textit{Finally, I checked the author’s proof towards the statement. I think the proofs need to be revised substantially as I found several flaws:}

    \textbf{Response:} We thank reviewer for pointing out these issues. We have provided a new proof of Theorem~\ref{thm:stats} for the new statistical impossibility statement. We will reply to each issue based on the related parts in the new proof.

    %Since we have proved a new stronger statement for statistical impossibility, most of following concerns no longer exist in the new proof of Theorem~\ref{thm:stats}. We still thank reviewer for pointing out these concerns and answer each sub-question for clarification.

    \begin{itemize}
        \item \textit{Page 32, Eqn (31) suggests that $\theta$ is fixed and the decision problems are reduced to $z$ from that point. But in latter argument $\theta$ is taken as unknown parameter.}

        \textbf{Response:} We agree that the inconsistent status of $\mtheta$ is a flaw in previous proof of Theorem~\ref{thm:stats}. Once we fix the true $\mtheta$ at the beginning (Eqn (31) of previous proof), we should take true $\mtheta$ as a known parameter in the rest of the proof. Specifically, in previous Eqn (33), the model of $\my$
        \begin{equation}
            \my = \theta \mx_z + \me
        \end{equation}
        should take $\theta$ as a known parameter. Because the distribution of observation $\my$ corresponds to the conditional probability $\bbP[\cdot | z_k(i)]$ in previous Eqn (31). When the true assignment $z_k(i)$ is known, the distribution of $\my$ should be known with true heterogeneity $\theta$ and true centroid $\mx_{z_k(i)}$. To avoid this flaw in the new proof, we use new notations $z^*_k, \mtheta^*_k, \tS^*_k$ for true parameters to better distinguish the known and unknown parameters. We also re-write the conditional probability as $\bbP[\cdot | z^*_k(i), z_k^*, \mtheta^*_k, \tS^*]$ for better expression.  We attach the related parts in the new proof in Appendix~\ref{sec:statprove1} here:

        \begin{quote}
            ``...
            To prove the minimax rate~\eqref{eq:minminmax}, it suffices to take an arbitrary $\tS^* \in  \tP_{\tS}(\gamma)$ wih $\gamma < -(K-1)$ and construct $(z^*_k, \mtheta^*_k)$ such that 
   \begin{equation}
       \inf_{\hat z_1} \bbE \left[ p\ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right]\geq 1.
   \end{equation}
   ..."
        \end{quote}

        \begin{quote}
            ``... we have 
   \begin{align}
        & \inf_{\hat z_1 } \bbE \left[ \ell(\hat z_1, z_1^*) | (z^*_k, \mtheta^*_k, \tS^*)  \right] \geq  \\
        & \frac{C}{r^3 |T_1^c|} \sum_{i \in T_1^c} \inf_{\hat z_1  (i)} \{ \bbP[\hat z_1(i) = 1| z_1^*(i) = 2, z_k^*, \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(i) = 2| z_1^*(i) = 1, z_k^*, \mtheta_k^*, \tS^*] \},
   \end{align}
    where $C$ is some positive constant, ..."
        \end{quote}

        \begin{quote}
            `` ...  The equivalent vector problem for assessing the summand in~\eqref{eq:inf_lower} is
\begin{equation}
\my=\theta^* \mx_{z}^*+\me,
\end{equation}
where $z\in \{1,2\}$ is an unknown parameter, $\theta^* \in \bbR_+$ is the given heterogeneity degree, $\mx_1^*,\mx_2^*\in\bbR^{p^{K-1}}$ are given centroids, and $\me\in\bbR^{p^{K-1}}$ consists of i.i.d.\ $N(0,\sigma^2)$ entries.  ..."
        \end{quote}

        \item \textit{Page 33, Line 9: let’s say the authors may want to make $\theta$ unknown and involve it in the decision problem. The authors formulate a hypothesis testing for model (33). My question is, is this really a hypotheses testing? Eqn. (30) involves both $z$ and $\theta$ as unknown parameters. Under null ($z = 1$), the distribution of $y$ is still unknown.}

        \textbf{Response:} This is a follow-up question of last comment. We agree that involving $\mtheta$ as an unknown parameter in the model of $\my$ and hypothesis testing is in appropriate. Following our last response, we consider the true heterogeneity $\mtheta^*$ as known in the model of $\my$. Hence, we have a valid simple hypothesis testing problem, in which the assignment function $z$ is the only unknown parameter in the test. Under the null hypothesis $(z = 1)$, the observation $\my = \theta^* \mx_1^* + \me$ follows multivariate Gaussian distribution with known mean vector $\theta^* \mx_1^*$ and covariance matrix $\sigma^2 \mI$. For clarification, we re-write the hypothesis testing with the model of $\my$. We attach the related parts in the new proof in Appendix~\ref{sec:statprove1} here:

        \begin{quote}
            ``...
            The equivalent vector problem for assessing the summand in~\eqref{eq:inf_lower} is
\begin{equation}
\my=\theta^* \mx_{z}^*+\me,
\end{equation}
where $z\in \{1,2\}$ is an unknown parameter, $\theta^* \in \bbR_+$ is the given heterogeneity degree, $\mx_1^*,\mx_2^*\in\bbR^{p^{K-1}}$ are given centroids, and $\me\in\bbR^{p^{K-1}}$ consists of i.i.d.\ $N(0,\sigma^2)$ entries.  Then, we consider the hypothesis testing under the model~\eqref{eq:z}:
\begin{equation}
 H_0: z = 1 \text{ and } \my = \theta^* \mx_1^* + \me \quad \leftrightarrow \quad H_1: z = 2 \text{ and } \my = \theta^* \mx_2^* + \me,
\end{equation}
...
            "
        \end{quote}

        % Based on last sub-question, we agree that the $\theta$ should be considered as a known parameter in model (33). The hypothesis testing from model (33) is equivalent to
        % \begin{equation}\label{eq:hypo_revise}
        %     H_0: \my = \theta \mx_1 + \me \quad \text{v.s.} \quad H_1: \my = \theta \mx_2 + \me,
        % \end{equation}
        % where $\theta, \mx_1, \mx_2$ are known parameters, $\me$ consists of i.i.d.\ $N(0, \sigma^2)$ entries. Now, the null and alternative distri-0bution of $\my$ is known, and the test~\eqref{eq:hypo_revise} is a valid hypothesis testing problem.

        \item \textit{A follow-up question is why could you apply Neyman-Pearson Lemma to lower bound the decision risks? Apparently this is not a simple testing, and the authors should provide the justification on why the presented “profiled MLE” works (technically) in determining the lower bound.}

        \textbf{Response:} The revised hypothesis testing in new Eqn~\eqref{eq:test} is a simple $(z = 1)$ versus simple $(z = 2)$ testing problem. Because the assignment function $z$ is the only unknown parameter in test. Therefore, we are able to apply Neyman-Pearson Lemma to our testing problem. By Neyman-Pearson Lemma, the likelihood ratio test, which coincides to the least square estimator of $z$ under our Gaussian model, leads to the minimal Type I + II error. As $\theta^*$ is a known parameter in the new proof, we no longer use the ``profile MLE" to determine the lower bound. We attach the related parts in the new proof in Appendix~\ref{sec:statprove1} here:

        \begin{quote}
            ``
            ... The hypothesis testing~\eqref{eq:test} is a simple versus simple testing, since the assignment $z$ is the only unknown parameter in the test. By Neyman-Pearson lemma, the likelihood ratio test is optimal with minimal Type I + II error. Under Gaussian model, the likelihood ratio test of \eqref{eq:test} is equivalent to the least square estimator $\hat z_{LS} = \argmin_{a = \{1,2\}} \onormSize{}{\my - \theta^* \mx_a^*}_F^2$.  ...
            "
        \end{quote}

        \item \textit{Although I don’t understand why $\hat \theta_{\rm MLE}$ is introduced here, I feel its calculation is incorrect. Does the authors consider the fact that there is a pre-specified parameter space for $\theta$ and it cannot take all values on $\bbR^+$ ?}

        \textbf{Response:} In the new proof, we no longer introduce the estimator $\hat \theta_{\rm MLE}$ and $\hat z_{\rm MLE}$ defined in the previous proof. Instead, we consider the least square estimator by Neyman-Pearson Lemma. In general, we agree that the MLE of $(z, \mtheta, \tS)$ should lie in the pre-specified parameter space $\tP$ to guarantee the identifiability. We emphasize this point in the formal definition of MLE~\eqref{eq:mle} in the main text. We attach the definition in Section~\ref{sec:limits} here:
        \begin{quote}
            ``
            We consider the Gaussian MLE, denoted as $(\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}})$, over the estimation space $\tP$, where 
\begin{equation}
    (\hat z_{\text{MLE}}, \hat \mtheta_{\text{MLE}}, \hat \tS_{\text{MLE}}) = \argmin_{ (z, \mtheta, \tS) \in \tP} \onormSize{}{\tY - \tX(z,\mtheta, \tS)}_F^2.
\end{equation}
...
            "
        \end{quote}

    \end{itemize}
    \end{enumerate}

\end{enumerate}

\newpage
\begin{center}
    \textbf{Point-by-point response to Reviewer 2}
\end{center}

\textit{I thank the authors for the signiﬁcant revision that clariﬁes the issues that I raised. I still suggest the authors to emphasise early that this paper focuses on dense regime. For instance, they could include sparsity in Table 1 and mention that Ahn et al (2018) proves exact recovery in the sparse regime (which this paper and few others in the list don't). This would provide a clear (unbiased) comparison of the results.}

\textbf{Response:} We thank reviewer for the reference. We have included the reference \cite{ahn2018hypergraph} in Table~\ref{tab:comp}:

\begin{quote}
    \begin{table*}[ht]
\resizebox{\textwidth}{!}{%
    \begin{tabular}{c|cccccc}
\hline
    & \cite{gao2018community}&  \cite{ahn2018hypergraph} &\cite{han2020exact}&  \cite{ghoshdastidar2017consistency} &\cite{ke2019community} & \textbf{Ours}\\
    \hline
      Allow tensors of arbitrary order & $\times$ & $\surd$&  $\surd$ & $\surd$ &$\surd$ & $\surd$  \\
        Allow degree heterogeneity &  $\surd$ & $\times$ & $\times$ & $\surd$ &$\surd$ & $\surd$ \\
        Singular-value gap-free clustering &  $\surd$ & $\surd$ & $\surd$ & $\times$ & $\times$ &$\surd$ \\
        Misclustering rate (for order $K^*$)& - & ${p^{-(K-1)} \alpha^{-1}_p}^{**}$ & $\exp(-p^{K/2})$ & $p^{-1}$& $p^{-2}$ &  $\exp(-p^{K/2})$\\
        Consider sparse observation & $\times$ & $\surd$ & $\times$ & $\times$ & $\times$ &$\times$\\
        \hline
    \end{tabular}
    }
    \caption{ Comparison between previous methods with our method. $^*$We list the result for order-K tensors with $K \geq 3$ and general number of communities $r = \tO(1)$. $^{**}$The parameter $\alpha = f(p) > 0$ denotes the sparsity level which is some function of dimension $p$. }
\end{table*}
\end{quote}

We have added a brief clarification in the related work, Section~\ref{sec:intro}:
\begin{quote}
    ``... Some works~\citep{ahn2018hypergraph} study the TBM with sparse observations, while, others~\citep{wang2019multiway, han2020exact} and our work focus on the dense regime. ..."
\end{quote}

We also added the reference in the extension to Bernoulli clustering, Section~\ref{subsec:ber}:
\begin{quote}
    ``The sparsity is often a popular feature in hypergraphs~\citep{florescu2016spectral,ke2019community, ahn2018hypergraph}.
    "
\end{quote}

\newpage
% \section{Notes}

% \textbf{Claim:} The statistical impossibility in dTBM can be easily obtained from the TBM paper.

%     \textcolor{blue}{ Note that ``easily obtained" is not ``directly obtained". See following arguments.}

%     Let $\tP_{\rm dTBM}$ and $\tP_{\rm TBM}$ denote the true parameter spaces for dTBM and TBM models. There is no belonging relation between $\tP_{\rm dTBM}$ and $\tP_{\rm TBM}$. Examples: if $\mtheta \neq \mathbf{1}$, then we have $(z, \mtheta, \tS) \in \tP_{\rm dTBM}$ but $(z, \mtheta, \tS) \notin \tP_{\rm TBM}$; if $\tS$ has $\onormSize{}{\mS_{a:}}_F = 0$ and $\mtheta = \mathbf{1}$, then we have $(z, \mtheta, \tS) \in \tP_{\rm TBM}$ but $(z, \mtheta, \tS) \notin \tP_{\rm dTBM}$. Also, there is no belonging relation between $\tP_{\rm dTBM}(\gamma) \coloneqq \tP_{\rm dTBM} \cap \{ \Delta_{\rm ang}^2 \lesssim p^{\gamma} \}$ and $\tP_{\rm TBM}(\gamma)  \coloneqq \tP_{\rm TBM} \cap \{ \Delta_{\rm Euc}^2 \lesssim p^{\gamma} \}$.

%     Hence, for any estimator $\hat z$, we can not infer the worst case error from one case to another case; i.e.,
%     \begin{equation}
%         \sup_{(z, \mtheta, \tS) \in \tP_{\rm dTBM}(\gamma)} p\ell(\hat z, z)  >  1 \quad \nRightarrow \text{and} \nLeftarrow \quad  \sup_{(z, \mtheta, \tS) \in \tP_{\rm TBM}(\gamma)} p\ell(\hat z, z) > 1.
%     \end{equation}
